---
title: "Accessing raw mass spectrometry data from a grover API"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Accessing raw mass spectrometry data from a grover API}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE
)
```

## Overview

This vignette outlines the `grover` package client-side functionality for accessing a raw mass spectrometry data repository hosted by grover API.
This functionality includes repository contents querying, file transfer, file information retrieval, sample information retrieval and raw data conversion.

To get started we can first load the `grover` package.

```{r setup}
library(grover)
```

## Example API

For this example we will run an example grover API provided by the package.
This will run as a background process to allow us to interact with the API without having to use an alterative R console.
To activate this run the following:

```{r grover_host}
grover_host <- grover(host = "127.0.0.1",
                     port = 8000,
                     auth = "1234",
                     repository = system.file('repository',
                                              package = 'grover'))

api <- groverAPI(grover_host,
                 background = TRUE,
                 log_dir = paste0(tempdir(),'/logs'))
```

For further details on hosting a grover API see the [Hosting a grover API](hosting.html) vignette.

```{r wait, echo=FALSE}
Sys.sleep(10)
```

## grover API host details

In order to access the API, we need to first provide the host details of the grover API.
There are two ways to do this.
The primary method is through the use of a configuration file that can then be parsed and specified when the API is activated.
This should be in YAML format and have the structure shown below:

```{r grover_client_config,echo=FALSE,comment=''}
config_file <- readLines(system.file('grover_client.yml',package = 'grover'))

cat(stringr::str_c(config_file,collapse = '\n'))
```

This specifies the host address, the port, and authentication key that matches the key with which the host has been configured.

The package contains an example file that we can parse here using the `readGrover()` function.

```{r grover_client_config_input}
grover_client <- readGrover(system.file('grover_client.yml',package = 'grover'))
```

This returns an S4 object of class `GroverClient` that contains the API host information.
The host information can be viewed by printing the object:

```{r print_grover_client}
print(grover_client)
```

The second method, and the method that will be used for this example, is to specify the host details directly using the `grover()` function like the following:

```{r grover_client}
grover_client <- grover(host = "127.0.0.1",
                     port = 8000,
                     auth = "1234")
```

This enables us to access the grover API hosting the example data repository included in the package.

Using this host information, we can first check that the API is running using the `extant()` function as shown below:

```{r extant}
extant(grover_client)
```

```{r}
message(api$is_alive())
```

```{r}
message(api$read_output())
```

```{r}
message(api$get_result())
```

## Querying the data repository contents

We can list the instruments available within the data reposiroty using the following: 

```{r listInstruments}
listInstruments(grover_client)
```

As can be seen above, there is a single instrument available named `r listInstruments(grover_client)`. 
To list the experiment directories available for this instrument, we can use the code below.

```{r listDirectories}
listDirectories(grover_client,'Thermo-Exactive')
```

This shows a single experiment data directory available.
We can then list the contents of this directory to identify the data files available:

```{r listFiles}
listFiles(grover_client,'Thermo-Exactive','Experiment_1')
```

We can see that there is a single raw data file available in this example repository.

## File information

File Information such as file size and creation date be retrieved. 
We can see this for the example file `QC01.raw` using the following:

```{r fileInfo}
fileInfo(grover_client,'Thermo-Exactive','Experiment_1','QC01.raw')
```

This can also be done directory wide when multiple files are available.

```{r directoryFileInfo}
directoryFileInfo(grover_client,'Thermo-Exactive','Experiment_1')
```

## Transfer files

Individual files can be transferred from the repository using `transferFile()`
, stipulating the instrument, experiment directory and file name.
The `outDir` argument allows us to declare where the file will be downloaded to.
In the example below, the file will be transfered to the current working directory.

```{r transferFile,eval=FALSE}
transferFile(grover_client,
             'Thermo-Exactive',
             'Experiment_1',
             'QC01.raw',
             outDir = '.')
```

Similarly, we can transfer an entire directory:

```{r transferDirectory,eval=FALSE}
transferDirectory(grover_client,
                  'Thermo-Exactive',
                  'Experiment_1',
                  outDir = '.')
```

## Sample information

Thermo `.raw` mass spectrometry data files contain sample meta information within the file headers.
This can be extracted and retrieved, in the form of a tibble, for a given file using:

```{r sampleInfo}
sampleInfo(grover_client,'Thermo-Exactive','Experiment_1','QC01.raw')
```

Similarly, the sample information for an entire experiment run can be retrieved with:

```{r runInfo}
runInfo(grover_client,'Thermo-Exactive','Experiment_1')
```

## Raw file conversion to .mzML format

With `grover` it is also possible retrieve `.mzML` format data files, converted from the `.raw` files.
This file conversion uses the command line tool [msconvert](http://proteowizard.sourceforge.net/tools.shtml), implemented in R by the [msconverteR](https://github.com/wilsontom/msconverteR) package.

### File conversion

To retrieve the example `.raw` file in `.mzML` format, the `convertFile()` function can be used.
This takes similar inputs as `transferFile()` shown previously.

```{r convertFile,eval=FALSE}
convertFile(grover_client,
            'Thermo-Exactive',
            'Experiment_1',
            'QC01.raw',
            outDir = '.')
```

```{r convertDirectory,eval=FALSE}
convertDirectory(grover_client,
            'Thermo-Exactive',
            'Experiment_1',
            outDir = '.')
```

### Conversion arguments

The `args` argument can be supplied to these conversion functions pass specific conversion criterial to `msconvert`.
The `grover` package contains a number of helper functions to simplify their use.
The available functions are listed below.

```{r conversion_functions,echo=FALSE,comment=''}
package_ls <- lsf.str("package:grover") 
conversion_functions <- package_ls[stringr::str_detect(package_ls,'conversion')]

cat(conversion_functions, sep = '\n')
```

Calling these functions return the appropriate string argument that is to be passed to msconvert.

```{r}
conversionArgsPeakPick()
```

Mutiple functions can also be combined.

```{r}
paste(conversionArgsPeakPick(),conversionArgsNegativeMode())
```

A full list of the available msconvert arguments can be found [here](http://proteowizard.sourceforge.net/tools.shtml).
The example below shows the use of the `conversionArgsPeakPick()` to retrieve a centroided data in `.mzML` format.

```{r convertFile_centroid,eval=FALSE}
convertFile(grover_client,
            'Thermo-Exactive',
            'Experiment_1',
            'QC01.raw',
            args = conversionArgsPeakPick(), 
            outDir = '.')
```

## Conclusion

To conclude, we can terminte the process running the example grover API instance.

```{r kill}
api$kill()
```